---
title: "Lobster Exploration"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(janitor)
library(here)
```

## SBC Lobsters

```{r}
lobsters <- read_csv(here("data","Lobster_Abundance_All_Years_20210412.csv"), na = c("-99999", "")) %>% 
  clean_names() %>%
  uncount(count)
```

### Find counts and mean sizes by site and year

```{r}
lobsters_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(lobster_count = n(), mean_length = mean(size_mm, na.rm = TRUE))
```

```{r}
ggplot(lobsters_summary, aes(x = year, y = lobster_count)) + geom_point(aes(color = site))
```

### Find the proportion of legal lobsters at each site for 2020

Legal lobster size in CA is 79.76 mm

```{r}
# Find counts of lobsters ovserced at each site that are above and below the legal limit
lobster_2020 <- lobsters %>% 
  filter(year == 2020) %>% 
  mutate(legal = case_when(
    size_mm >= 79.76 ~ "yes",
    size_mm < 79.76 ~"no"
  )) %>%
  group_by(site, legal) %>%
  summarize(count = n()) %>% 
```

Create a stacked column graph that shows the *proportion* of legal and non-legal lobsters at each site. \*\*Hint: create a stacked column graph with `geom_col()`, then add the argument `position = "fill"` to convert from a graph of absolute counts to proportions.

```{r}
ggplot(lobster_2020, aes(x = site, y = count, color = legal, fill = legal)) + 
  geom_col(position = "fill")
```

IVEE and NAPL had the largest proportion of legal lobsters.

## Random lobster wrangling

```{r}
# filter() practice
# Create and store a subset that only contains lobsters from sites “IVEE”, “CARP” and “NAPL”. Check your output data frame to ensure that only those three sites exist.
ex_a <- lobsters %>% 
  filter(site %in% c("IVEE", "CARP", "NAPL"))

```

```{r}
# Create a subset that only contains lobsters observed in August
ex_b <- lobsters %>% 
  filter(month == 8)
```

```{r}
# Create a subset with lobsters at Arroyo Quemado (AQUE) OR with a carapace length greater than 70 mm.
ex_c <- lobsters %>% 
  filter(site == "AQUE" | size_mm > 70)
```

```{r}
# Create a subset that does NOT include observations from Naples Reef (NAPL)
ex_d <- lobsters %>% 
  filter(site != "NAPL")
```

group_by() %\>% summarize() practice

```{r}
# Find the mean and standard deviation of lobster carapace length, grouped by site.
ex_e <- lobsters %>% 
  group_by(site) %>% 
  summarize(mean_length = mean(size_mm, na.rm = TRUE),
            sd_length = sd(size_mm, na.rm = TRUE))
```

```{r}
# Find the maximum carapace length by site and month.
ex_f <- lobsters %>% 
  group_by(site, month) %>% 
  summarize(max_length = max(size_mm, na.rm = TRUE))
```

mutate() practice

```{r}
# Add a new column that contains lobster carapace length converted to centimeters. Check output.
ex_g <- lobsters %>% 
  mutate(length_cm = size_mm / 10)
```

```{r}
# Update the site column to all lowercase. Check output.
ex_h <- lobsters %>% 
  mutate(site = tolower(site))
```

```{r}
# Convert the area column to a character (not sure why you’d want to do this, but try it anyway). Check output.
ex_i <- lobsters %>% 
  mutate(area = as.character(area))
```

case_when() practice

```{r}
# Use case_when() to add a new column called size_bin that contains “small” if carapace size is <= 70 mm, or “large” if it is greater than 70 mm. Check output.
ex_j <- lobsters %>% 
  mutate(size_bin = case_when(
    size_mm <= 70 ~ "small",
    size_mm > 70 ~ "large"
  ))
```

```{r}
# Use case_when() to add a new column called designation that contains “MPA” if the site is “IVEE” or “NAPL”, and “not MPA” for all other outcomes.
ex_k <- lobsters %>% 
  mutate(designation = case_when(
    site %in% c("IVEE", "NAPL") ~ "MPA",
    TRUE ~ "not MPA"
  ))
```

\
